Formalization of Abstract Rewriting

### 2025-02-23 Relations/Paper/Implications.tex
call update

We are interested in seeing which combination of properties a relation requires in order to be complete (that is, $\CR$ and $\SN$). 
The following tables show either that a given combination is complete, or the appropriate counterexample which demonstrates why 
those properties do not provide a property higher on the taxonomical hierarchy.

@@

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\begin{table}[h!]
    \centering
    \caption{Local implications}
    \begin{tabular}{|>{\columncolor{gray!30}}l|c|c|c|c|c|}
    \hline
    \rowcolor{gray!30}     & $\WM$         & $\WN$         & $\SM$             & $\SMandWN$        & $\SN$ \\
    \hline
    $\UNto$ &  CE-\ref{CE:4}, CE-\ref{CE:12}     & CE-\ref{CE:4}, CE-\ref{CE:7}       & CE-\ref{CE:4} , CE-\ref{CE:9}          & CE-\ref{CE:4}, CE-\ref{CE:4}            & $\CR$\footnotemark[1] , $\SN$ \\
    \hline
    $\NP$ & CE-\ref{CE:10},CE-\ref{CE:9}      & $\CR$, CE-\ref{CE:7}    & CE-\ref{CE:11} , CE-\ref{CE:9}          & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\MP$ & $\CR$ , CE-\ref{CE:9}    & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[2] , CE-\ref{CE:9} & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\CR$   & $\CR$, CE-\ref{CE:9}     & $\CR$, CE-\ref{CE:7}     & $\CR$, CE-\ref{CE:9}        & $\CR$, $\SN$     & $\CR$ , $\SN$ \\
    \hline

    \end{tabular}
\end{table}
\begin{table}[h!]
    \centering
    \caption{Global implications}
    \begin{tabular}{|>{\columncolor{gray!30}}l|c|c|c|c|c|}
    \hline
    \rowcolor{gray!30}     & $\WM$         & $\WN$         & $\SM$             & $\SMandWN$        & $\SN$ \\
    \hline
    $\UNto$ &  CE-\ref{CE:2}, CE-\ref{CE:9}     & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[1] , CE-\ref{CE:9}          & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\NP$ & CE-\ref{CE:3} , CE-\ref{CE:9}      & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[1] , CE-\ref{CE:9}          & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\MP$ & $\CR$ , CE-\ref{CE:9}    & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[2] , CE-\ref{CE:9} & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\CR$   & $\CR$, CE-\ref{CE:9}     & $\CR$, CE-\ref{CE:7}     & $\CR$, CE-\ref{CE:9}        & $\CR$, $\SN$      & $\CR$ , $\SN$ \\
    \hline

    \end{tabular}
\end{table}
\footnotetext[1]{This implication also requires global $\WCR$.}
\footnotetext[2]{This implication also requires either global $\WCR$ or the classical property required to go from 
$\SN \to \WN$ or $\SM \to \WM$.}
\renewcommand*{\thefootnote}{\arabic{footnote}}

@@

\sacomment{Need to update the tables with appropriate symbols for global and local. Colour for first row and column?}

\sacomment{Rewrite the below to reflect changes to the table.}
The tables show a number of changes when moving from properties holding locally to globally. 
Globally $\WN \land \UNto \implies \CR$ holds whereas it does not locally. Similarly, $\SM \land \WN \implies \SN$ holds 
globally but not locally.

Theorem 1.2.2 ii in \terese is the claim that $\WN \land \UN \implies \CR$ holds, and we formalize this in

\verb|ii : R isWN × R isUN → R isCR| \footnotemark[3]

We have modestly generalized this theorem by showing that $\WN \land \UNto \implies \CR$ holds (and indeed this is the basis for our proof of the theorem). 
Counterexample \sacomment{REF HERE} shows why this implication cannot hold when the properties are only local. Our proof is in the function:

\verb|UN→∧WN→CR : R isUN→ → R isWN → R isCR| \footnotemark[1]

To show that $\SM \land \WN \implies \SN$ we built on our proof that locally $\SM \land \WN \land \NP \implies \SN$. If $\WN$ applies 
globally then our proof no longer requires the property $\NP$. This progression of proofs can be seen in the functions:

@@

\verb|isWN∧isSM→isSN : R isWN → R isSM → R isSN| \footnotemark[1]

that $\SN$ implies $\SM$, and it is from this implication 
that we produce a new generalization of Newman's lemma. Every relation with the properties 
$\SM$ and $\WCR$ has the property of $\CR$. This holds true both locally and globally as we show in:

\verb|LocalNewmansLemmaRecurrent : R isWCR → SM ⊆ CR| \footnotemark[2]

and \verb|GlobalNewmansLemmaRecurrent : R isWCR → R isSM → R isCR| \footnotemark[2]

\sacomment{We should remember that If global WCR and local SM implies local CR, then if we also have local WN, we get local SN.}

>>>>

We are interested in seeing which combination of properties a relation requires in order to be complete (that is, $\gCR$ and $\gSN$). 
The following tables show either that a given combination is complete, or the appropriate counterexample which demonstrates why 
those properties do not provide a property higher on the taxonomical hierarchy.

@@

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\begin{table}[h!]
    \centering
    \caption{Local implications}
    \begin{tabular}{|>{\columncolor{gray!30}}l|c|c|c|c|c|}
    \hline
    \rowcolor{gray!30}     & $\WM$         & $\WN$         & $\SM$             & $\SMandWN$        & $\SN$ \\
    \hline
    $\UNto$ &  CE-\ref{CE:4}, CE-\ref{CE:12}     & CE-\ref{CE:4}, CE-\ref{CE:7}       & $\CR$\footnotemark[1] , CE-\ref{CE:9}          & $\CR$\footnotemark[1] , CE-\ref{CE:4}            & $\CR$\footnotemark[1] , $\SN$ \\
    \hline
    $\NP$ & CE-\ref{CE:10},CE-\ref{CE:9}      & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[1]  , CE-\ref{CE:9}          & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\MP$ & $\CR$ , CE-\ref{CE:9}    & $\CR$, CE-\ref{CE:7}    & $\CR$\footnotemark[2] , CE-\ref{CE:9} & $\CR$, $\SN$      & $\CR$\footnotemark[2] , $\SN$ \\
    \hline
    $\CR$   & $\CR$, CE-\ref{CE:9}     & $\CR$, CE-\ref{CE:7}     & $\CR$, CE-\ref{CE:9}        & $\CR$, $\SN$     & $\CR$ , $\SN$ \\
    \hline

    \end{tabular}
\end{table}
\begin{table}[h!]
    \centering
    \caption{Global implications}
    \begin{tabular}{|>{\columncolor{gray!30}}l|c|c|c|c|c|}
    \hline
    \rowcolor{gray!30}     & $\gWM$         & $\gWN$         & $\gSM$             & $\gSMandWN$        & $\gSN$ \\
    \hline
    $\gUNto$ &  CE-\ref{CE:2}, CE-\ref{CE:9}     & $\gCR$, CE-\ref{CE:7}    & $\gCR$\footnotemark[1] , CE-\ref{CE:9}          & $\gCR$, $\gSN$      & $\gCR$\footnotemark[2] , $\gSN$ \\
    \hline
    $\gNP$ & CE-\ref{CE:3} , CE-\ref{CE:9}      & $\gCR$, CE-\ref{CE:7}    & $\gCR$\footnotemark[1] , CE-\ref{CE:9}          & $\gCR$, $\gSN$      & $\gCR$\footnotemark[2] , $\gSN$ \\
    \hline
    $\gMP$ & $\gCR$ , CE-\ref{CE:9}    & $\gCR$, CE-\ref{CE:7}    & $\gCR$\footnotemark[2] , CE-\ref{CE:9} & $\gCR$, $\gSN$    & $\gCR$\footnotemark[2] , $\gSN$ \\
    \hline
    $\gCR$   & $\gCR$, CE-\ref{CE:9}     & $\gCR$, CE-\ref{CE:7}     & $\gCR$, CE-\ref{CE:9}        & $\gCR$, $\gSN$      & $\gCR$ , $\gSN$ \\
    \hline

    \end{tabular}
\end{table}
\footnotetext[1]{This implication also requires $\gWCR$.}
\footnotetext[2]{This implication also requires either $\gWCR$ or the classical property required to go from 
$\SN \to \WN$ or $\SM \to \WM$.}
\renewcommand*{\thefootnote}{\arabic{footnote}}

@@

\subsubsection{Generalized Newman's Lemma}\label{subsec:newnewman}
One interesting outcome from investigating the normalization taxonomy was a generalization of Newman's Lemma.

\begin{proposition}
    $\gWCR \land \SM \implies \CR$
\end{proposition}
\begin{proof}
    See, \verb|LocalNewmansLemmaRecurrent : R isWCR → SM ⊆ CR|
\end{proof}

This proof follows the same steps as the second proof of Newman's Lemma found in \terese, but 
uses the generalization $\SM$ of $\SN$.

% The tables show a number of changes when moving from properties holding locally to globally. 
% Globally $\WN \land \UNto \implies \CR$ holds whereas it does not locally. Similarly, $\SM \land \WN \implies \SN$ holds 
% globally but not locally. 
\subsection{Relation between $\SM$, $\WN$, and $\SN$}
One key takeaway from the tables is that we obtain completeness when we have the properties $\SM$, $\WN$, and $\WN$. We also 
obtain completeness when we have $\gWN$ and $\gSM$. 
To show that $\gSM \land \gWN \implies \gSN$ we built on our proof that $\SM \land \WN \land \NP \implies \SN$. If $\gWN$ holds 
then our proof no longer requires the property $\NP$. This progression of proofs can be seen in the functions:

@@

\verb|isWN∧isSM→isSN : R isWN → R isSM → R isSN|

\sacomment{Are there any other features of the table that we would like to draw out?}
\sacomment{We should remember that If global WCR and local SM implies local CR, then if we also have local WN, we get local SN.}
